---
title: "Trumps' Economy Adjust Poll Numbers"
author: "Salim Damerdji"
output: html_document
---

```{r, include=F}
library(tidyverse)
```

The economy is great, but Trump's approval ratings are low by historical standards. What would the polls say if the economy were average? Let's find out.

I use two datasets for approval ratings: 538's approval ratings for Trump, and Gallup's approval ratings for presidents before Trump (obtiained from the American Presidency Project.) To use the latter dataset, I had to manually copy/paste all cells to one sheet. Ugh!

Data from:
APP:
https://www.presidency.ucsb.edu/statistics/data/presidential-job-approval
538:
https://github.com/fivethirtyeight/data/tree/master/trump-approval-ratings
GDP:
https://datahub.io/core/gdp-us
Employment:
https://datahub.io/core/employment-us#resource-aat1

# Data on Approval Ratings
```{r}
before_trump <- as.tibble(read.csv(file = "./potus_polls.csv"))
trump <- as.tibble(read.csv(file = "./trump_polls.csv"))
```

```{r}
#   filter(grade %in% c("A+","A","A-","B+","B"))
(trump <- trump  %>%
   mutate(startdate = as.Date(startdate, format = "%m/%d/%Y"),
          enddate = as.Date(enddate, format = "%m/%d/%Y")) %>%
   select(president, startdate, enddate, approve, disapprove))
  
(before_trump <- before_trump %>%
    mutate(startdate = as.Date(Start.Date, format = "%m/%d/%Y"), 
         enddate = as.Date(End.Date, format = "%m/%d/%Y")) %>%
    rename(president = ï..President,
         approve = Approving,
         disapprove = Disapproving) %>%
    select(-Unsure.NoData, -End.Date, -Start.Date))

(approval <- as.tibble(merge(trump, before_trump, all = TRUE)))

approval %>% 
  count(president) %>%
  arrange(desc(n))

apply(approval, 2, function(x) any(is.na(x)))

range(approval$startdate)
```

# Data on the Economy
## GDP Data
```{r}
(gdp <- as.tibble(read.csv(file = "./gdp_by_quarter.csv")) %>%
    mutate(date = as.Date(date, format = "%m/%d/%Y")))

(gdp_year <- as.tibble(read.csv(file = "./gdp_by_year.csv")) %>%
    mutate(date = as.Date(paste(date, 1, 1, sep = "-"))) %>%
    filter(date < min(gdp$date)))

(gdp <- as.tibble(merge(gdp, gdp_year, all = TRUE)) %>%
    filter(date > min(approval$startdate)))
```
## Employment Data
```{r}
(employment_year <- as.tibble(read.csv(file = "./employment_by_year.csv")))
(employment_quarter <- as.tibble(read.csv(file = "./employment_by_quarter.csv")))

(employment_year <- employment_year %>%
    mutate(year = as.Date(paste(year, 1, 1, sep = "-"))) %>%
    rename(date = year) %>%
    select(date, unemployed_percent))

(employment_quarter <- employment_quarter %>% 
    gather("month","unemployed_percent", 2:13, na.rm = TRUE))

(employment_quarter <- employment_quarter %>%
    mutate(date = as.Date(paste(match(employment_quarter$month, month.abb),"01", employment_quarter$ï..Year, sep="-"),
                          format = "%m-%d-%Y")) %>%
    select(date, unemployed_percent))

employment_year <- employment_year %>%
  filter(date < min(employment_quarter$date))

range(employment_quarter$date)
range(employment_year$date)
(employment <- as.tibble(merge(employment_year, employment_quarter, all = TRUE)))
```


```{r}
ggplot(data = approval) +
  geom_line(aes(x = startdate, y = (approve - disapprove)), color = "black") +
  geom_line(data = employment, aes(x = date, y = (unemployed_percent)), color = "blue") +
  geom_line(data = gdp, aes(x = date, y = change.chained), color = "red") +
  scale_x_date(date_minor_breaks = "1 year") + 
  xlab("Date") + 
  ylab("Approval Rating, GDP, Employment")
```
