---
title: "How Popular is Trump ?"
output: html_document
---

Notes to self:

- NAs are being introduced somehow before get_unemployment() or get_gdp()


- I don't have GDP data after 2017. This is really important since I'm using only prior gdp data


- Figure out when gdp data gets collected


- Add variable for time since last inauguration. Apparently presidents get a bump when running again


- Smooth polling data to get rid of noise


```{r, include=F}
library(tidyverse)
library(lubridate)
```

```{r}
get_tibble <- function(filename, date_format){
  return(as_tibble(read.csv(paste0("../data/cleandata/", filename, ".csv"))) %>%
            mutate(date = as.Date(date, format = date_format)))
}

approval <- get_tibble("approval", "%Y-%m-%d") %>% 
  arrange(date)
gdp <- get_tibble("gdp", "%Y-%m-%d")
employment <- get_tibble('employment', "%Y-%m-%d")
cpi <- get_tibble('cpi', "%Y-%m-%d")
exports <- get_tibble('export', "%Y-%m-%d")
housing <- get_tibble('housing', "%Y-%m-%d")
term_dates <- get_tibble('inauguration', "%m-%d-%Y") %>%
  rename(term_start=date)
```

```{r}
approval %>%
  group_by(president) %>%
  count() %>%
  arrange(desc(n))
```
## Feature Engineering

#### Days as president

```{r}
presidency_start <- term_dates %>%
  group_by(president) %>%
  summarise(pres_start = min(term_start))

approval <- left_join(approval, presidency_start, by = 'president') %>%
  mutate(days_as_pres = as.double(date - pres_start)) %>%
  select(-pres_start)
```

#### Days into term
```{r}

term_dates_stack <- term_dates$term_start
# This functions assumes term_dates and approval are sorted such that earlier dates are on top
# This function also will remove all but one observation from term_dates_stack
# Using a stack is 12x faster than brute force with for loop, and 10x faster than brute force with lapply
get_days_into_term <- function(row_num){
  p_date <- approval$date[row_num]
  if (length(term_dates_stack)==1 || (p_date>term_dates_stack[1] && p_date<term_dates_stack[2])){
    return(p_date - term_dates_stack[1])
  } else if (p_date > term_dates_stack[2]) {
    term_dates_stack <<- term_dates_stack[-1]
    return(get_days_into_term(row_num))
  } else {
    return(NA)
  }
}
approval[,'days_into_term'] <- sapply(1:nrow(approval), get_days_into_term) 

```

#### Years of Growth
```{r, include=F}

gdp <- mutate(gdp, is_gdp_up = sapply(gdp$gdp_change, function(change) if_else(change > 0, true = 1, false = 0)))
gdp[,'qrtrs_of_growth'] <- NA

gdp
cum_change <- function(row){
  if (row == 1) {
    gdp[row, "qrtrs_of_growth"] <<- 0
  }
  else if (gdp$is_gdp_up[row] == 1 && gdp$is_gdp_up[row-1] >= 1) {
    gdp[row, "qrtrs_of_growth"] <<- gdp$qrtrs_of_growth[row-1] + 1
  }
  else if (gdp$is_gdp_up[row] == 0 && gdp$is_gdp_up[row-1] <= 0) {
    gdp[row, "qrtrs_of_growth"] <<- gdp$qrtrs_of_growth[row-1] - 1
  }
  else {
    gdp[row, "qrtrs_of_growth"] <<- gdp$is_gdp_up[row]
  }
}

sapply(1:nrow(gdp), cum_change)
```

```{r}
gdp %>% filter(

```

Note to self: the above features absolutely introduce collinearity.

```{r}

range(gdp$date)
range(employment$date)
range(cpi$date)
range(housing$date)
range(exports$date)
range(approval$date)


get_gdp <- function(poll_day){
  return((filter(gdp, date < poll_day) %>%
           filter(date == max(date)) %>%
           select(gdp_change) %>%
           as.numeric())[1]) 
}

get_unemployment <- function(poll_day){
  return((filter(unemployment, date < poll_day) %>%
           filter(date == max(date)) %>%
           select(unemployed_percent) %>%
           as.numeric())[1]) 
}
```

```{r}
approval <- approval %>%
  mutate(gdp = sapply(X = approval$startdate, FUN = get_gdp),
         unemployment = sapply(X = approval$startdate, FUN = get_unemployment))

# Deal with NAs later
approval <- approval[complete.cases(approval), ]
nrow(approval)

(approval)

smoother <- approxfun(x = approval$startdate, y = approval$approve)

approval_smoothed <- approval %>% 
  select(-enddate, -disapprove) %>%
  mutate(poll_smoothed = smoother(startdate))

model <- lm(poll_smoothed ~ days_as_pres + gdp + unemployment + president, data = approval_smoothed, x = TRUE)
summary(model)

ggplot(data = as.data.frame(model$residuals)) +
  geom_point(aes(x = approval$startdate, y = model$residuals))

predicted_approval <- data.frame(pred= predict(model, approval_smoothed), x = approval_smoothed$startdate)

(approval_smoothed)
ggplot(data = approval_smoothed, aes(x = startdate, y = poll_smoothed)) +
  geom_smooth(method = "lm", formula = approval_smoothed$poll_smoothed ~ days_as_pres + gdp + unemployment + president)

ggplot(approval_smoothed) +
  geom_point(aes(x = startdate, y = poll_smoothed), color = "steelblue", alpha = .1) +
  geom_line(aes(x = predicted_approval$x, y = predicted_approval$pred), color = "black")
  
```


Normal unemployment rate is 5%

```{r}

sqrt(mean((summary(model))$residuals^2))
avg_unemployment_per_year <- unemployment %>%
  group_by(floor_date(date, "year")) %>%
  summarise(avg = mean(unemployed_percent))

avg_gdp_growth_per_year <- gdp %>%
  group_by(floor_date(date, "year")) %>%
  summarise(avg = mean(gdp$gdp_change))


(normal_unemployment <- mean(avg_unemployment_per_year$avg))
(normal_gdp <- mean(avg_gdp_growth_per_year$avg))

squo_unemployment <- (filter(unemployment, date == max(date)))$unemployed_percent
squo_gdp <- filter(gdp, date == max(date))$gdp_change

trump_econ_bump <- (normal_unemployment - squo_unemployment)* 3.39 + (squo_gdp - normal_gdp) * .007
(trump_recent <- approval %>%
    filter(startdate == max(startdate)) %>%
    select(approve, disapprove) %>%
    summary()) 

40 - trump_econ_bump

```